### 2023-05-07 manuscript/manuscript.qmd
Use separate profiles + word count

---
title: "Amicable Contempt: A Conceptual Framework for Understanding International NGO Behavior in the Era of Closing Civic Space"
short-title: "Amicable Contempt"
author:
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
bibliography: references.bib
---

Lorem ipsum dolor sit amet, consectetur adipisicing elit [@Mayer:2010], sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

# References