### 2023-05-07 manuscript/appendix.qmd
Add appendix placeholder

---
title: 'Online appendix for "Amicable Contempt: A Conceptual Framework for Understanding International NGO Behavior in the Era of Closing Civic Space"'
short-title: 'Online appendix for "Amicable Contempt"'
author:
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
bibliography: references.bib
crossref: 
  tbl-title: Appendix Table
  fig-title: Appendix Figure
---

Here's the appendix

```{r warning=FALSE, message=FALSE}
library(gt)

mtcars |>
  gt()
```