### 2023-05-07 notebook/paper.qmd
Generate paper links automatically

---
title: Paper
toc: false
page-layout: full
---

```{r include=FALSE}
library(dplyr)
library(glue)

links_to_include <- tribble(
  ~title, ~path, 
    ~icon, ~btn_class, ~target,
  "HTML", "/manuscript/output/manuscript.html", 
    '<i class="fa-solid fa-file-lines"></i>', "btn-success", "manuscript",
  "HTML <small>(new window)</small>", "/manuscript/output/manuscript.html", 
    '<i class="fa-solid fa-arrow-up-right-from-square"></i>', "btn-success", "_blank",
  "PDF", "/manuscript/output/manuscript.pdf", 
    '<i class="fa-solid fa-file-pdf"></i>', "btn-danger", "manuscript",
  "PDF <small>(new window)</small>", "/manuscript/output/manuscript.pdf", 
    '<i class="fa-solid fa-file-arrow-down"></i>', "btn-danger", "_blank",
  "Manuscripty PDF", "/manuscript/output-ms/manuscript.pdf", 
    '<i class="fa-solid fa-file-pdf"></i>', "btn-warning", "manuscript",
  "Manuscript PDF <small>(new window)</small>", "/manuscript/output-ms/manuscript.pdf", 
    '<i class="fa-solid fa-file-arrow-down"></i>', "btn-warning", "_blank"
) %>% 
  mutate(nice_link = glue(
    '<a class="btn [btn_class] btn-sm mb-1" href="[path]" target="[target]">',
    '[icon] &ensp;[title]',
    '</a>', .open = "[", .close = "]"
  ))

links <- paste(links_to_include$nice_link, collapse = "\n")
```

```{=html}
<p class="text-center">
  `r links`
</p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:90%;width:100%;border:2px solid #ccc;" height="90%" width="100%" name="manuscript"></iframe>
```