### 2023-05-09 notebook/descriptive.qmd
Add where NGOs work section

---
title: "Descriptive stuff"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "ragg_png")

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(sf)
library(countrycode)
library(scales)
library(gt)
library(skimr)
library(ggalluvial)
library(here)

tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(survey_orgs, survey_countries, survey_all))
tar_load(c(world_map))

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
set_annotation_fonts()

opts_int <- function(x, ...) {
  x %>% 
    opt_interactive(use_compact_mode = TRUE, use_highlight = TRUE, ...)
}

opts_theme <- function(x) {
  x %>% 
    opt_table_font(font = "Fira Sans")
}
```

# General respondent details

## How many NGOs responded?

```{r summary-n-respondents}
#| code-fold: show
nrow(survey_orgs)
```

## How many respondents answered questions for more than one country?

```{r summary-n-multiple-countries}
#| code-fold: show
survey_countries %>% 
  summarize(more_than_1 = sum(loop.number > 1),
            prop = more_than_1 / nrow(survey_orgs))
```

## Who in the organization responded to the survey?

```{r plot-respondents-who, fig.width=6, fig.height=3}
df_plot_respondents <- survey_orgs %>%
  group_by(Q2.3) %>%
  summarise(num = n()) %>%
  arrange(num) %>% 
  mutate(question = fct_inorder(Q2.3))

df_plot_respondents %>% 
  ggplot(aes(x = num, y = question)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0),
                     sec.axis = sec_axis(~ . / sum(df_plot_respondents$num),
                                         labels = label_percent())) +
  labs(x = "Respondents", y = NULL,
       title = "Who filled out the survey?",
       subtitle = "Q2.3: What is your position in your organization?") +
  theme_ingo()
```

Lots of others. Who are the others?

```{r tab-who-responded-others}
survey_orgs %>%
  filter(!is.na(Q2.3_TEXT)) %>%
  mutate(`Position in organization` = str_to_title(Q2.3_TEXT)) %>%
  group_by(`Position in organization`) %>%
  summarise(Count = n()) %>%
  mutate(Proportion = Count / sum(Count)) %>%
  arrange(desc(Count)) %>% 
  gt() %>% 
  fmt_percent(columns = Proportion, decimals = 1) %>% 
  opts_int() %>% opts_theme()
```

# Where are NGOs based?

## How are these NGOs distributed by HQ?

```{r tab-hq-countries}
df_hq_countries <- survey_orgs %>% 
  group_by(Q2.2_iso3) %>% 
  summarize(num_ngos = n()) %>% 
  ungroup() %>% 
  mutate(num_ngos_ceiling = ifelse(num_ngos >= 30, 30, num_ngos),
         ngo_presence = ifelse(num_ngos >= 1, "At least one NGO respondent", NA),
         prop = num_ngos / sum(num_ngos, na.rm = TRUE)) %>% 
  mutate(country_name = countrycode(Q2.2_iso3, "iso3c", "country.name")) %>%
  arrange(desc(num_ngos))

df_hq_countries %>% 
  select(country_name, Q2.2_iso3, num_ngos, prop) %>%
  gt() %>% 
  cols_label(
    country_name = "Country name",
    Q2.2_iso3 = "ISO3",
    num_ngos = "NGOs",
    prop = "Proportion"
  ) %>% 
  fmt_percent(columns = prop, decimals = 1) %>% 
  opts_int() %>% opts_theme()
```

## Number of unique HQ countries

```{r summary-num-hq-countries}
#| code-fold: show
df_hq_countries %>% 
  summarize(total = sum(num_ngos >= 1, na.rm = TRUE))
```

## Regional distribution of HQ countries

```{r tab-hq-regions}
df_hq_regions <- df_hq_countries %>%
  mutate(region = countrycode(Q2.2_iso3, "iso3c", "continent"),
         region = ifelse(Q2.2_iso3 == "TWN", "Asia", region)) %>%
  mutate(region = recode(region, Oceania = "Asia &\nOceania", Asia = "Asia &\nOceania")) %>%
  group_by(region) %>%
  summarise(num = sum(num_ngos)) %>% ungroup() %>%
  mutate(prop = num / sum(num)) %>%
  arrange(desc(num)) %>%
  mutate(region = fct_rev(fct_inorder(region)))

df_hq_regions %>% 
  gt() %>% 
  cols_label(
    region = "Region",
    num = "NGOs",
    prop = "Proportion"
  ) %>% 
  fmt_percent(columns = prop, decimals = 1) %>% 
  cols_align(align = "left", columns = region) %>% 
  opts_theme() %>% 
  opts_int(use_pagination = FALSE, use_pagination_info = FALSE)
```

```{r plot-hq-regions, fig.width=6, fig.height=3}
ggplot(df_hq_regions, aes(x = num, y = region)) +
  geom_pointrange(aes(xmin = 0, xmax = num)) +
  scale_x_continuous(sec.axis = sec_axis(~ . / sum(df_hq_regions$num),
                                         labels = label_percent())) +
  labs(x = "Respondents based in region", y = NULL,
       title = "Region of headquarters",
       subtitle = "Q2.2: Where is your organization's headquarters? (summarized by region)") +
  theme_ingo() + 
  theme(panel.grid.major.y = element_blank())
```

## Responses per country (30 NGO ceiling)

```{r plot-map-respondents-gradient}
map_hq <- world_map %>% 
  left_join(df_hq_countries, by = join_by(ISO_A3 == Q2.2_iso3))

ggplot() +
  geom_sf(data = map_hq, aes(fill = num_ngos_ceiling),
          linewidth = 0.1, color = "grey30") +
  scale_fill_binned(low = clrs$Peach[1], high = clrs$Peach[7], na.value = "grey95",
                    labels = c(1, 10, 20, "30+"), breaks = c(1, 10, 20, 30),
                    name = "Number of NGOs: ",
                    guide = guide_colorsteps(barwidth = 7, barheight = 0.4,
                                             title.vjust = 1)) +
  coord_sf(crs = "+proj=robin") +
  theme_ingo_map()
```

## Countries with at least one response

```{r plot-map-respondents-binary}
ggplot() +
  geom_sf(data = map_hq, aes(fill = ngo_presence),
          linewidth = 0.1, color = "grey30") +
  scale_fill_manual(values = c(clrs$Peach[7], "grey95"), 
                    na.value = "grey95", na.translate = FALSE,
                    name = NULL) +
  coord_sf(crs = "+proj=robin") +
  theme_ingo_map()
```

# Where do NGOs work?

## How are these NGOs distributed by target country?

```{r tab-target-countries}
possible_countries <- tibble(ISO3 = unique(as.character(world_map$ISO_A3))) %>% 
  filter(ISO3 != "-99")

df_work_countries_all <- survey_orgs %>%
  unnest(Q2.5_iso3) %>%
  group_by(Q2.5_iso3) %>%
  summarise(num_ngos = n()) %>%
  right_join(possible_countries, by = join_by(Q2.5_iso3 == ISO3)) %>% 
  mutate(prop = num_ngos / sum(num_ngos, na.rm = TRUE),
         country_name = countrycode(Q2.5_iso3, "iso3c", "country.name",
                                    custom_match = c("XKX" = "Kosovo")),
         presence = num_ngos >= 1) %>%
  arrange(desc(num_ngos))

df_work_countries_answered <- survey_all %>%
  group_by(Q4.1_iso3) %>%
  summarise(num = n())

df_work_countries_all %>% 
  select(country_name, Q2.5_iso3, num_ngos, prop) %>%
  gt() %>% 
  cols_label(
    country_name = "Country name",
    Q2.5_iso3 = "ISO3",
    num_ngos = "NGOs",
    prop = "Proportion"
  ) %>% 
  fmt_percent(columns = prop, decimals = 1) %>% 
  opts_int() %>% opts_theme()
```

## Number of unique target countries

Respondents indicated working in this many unique countries:

```{r summary-num-target-countries}
#| code-fold: show
df_work_countries_all %>% 
  summarize(total = sum(num_ngos >= 1, na.rm = TRUE))
```

```{r summary-num-countries-answered}
df_work_countries_answered %>% 
  nrow()
```

## Countries per response

```{r tab-countries-per-response}
countries_per_response <- survey_orgs %>%
  unnest(Q2.5_iso3) %>%
  group_by(clean.id) %>%
  summarise(num = n())

skim_without_charts(countries_per_response$num) %>% 
  select(-c(skim_type, skim_variable, complete_rate)) %>% 
  rename_with(~str_remove(., "numeric\\.")) %>% 
  gt() %>% 
  opts_theme() %>% 
  opt_horizontal_padding(scale = 2)
```

```{r plot-countries-per-response, fig.width=6, fig.height=3}
ggplot(countries_per_response, aes(x = num)) +
  geom_histogram(binwidth = 2, color = "white", linewidth = 0.1) + 
  labs(x = "Number of countries selected", y = "Frequency") +
  theme_ingo()
```

## Flows between countries selected

```{r plot-continents-marked}
continents_marked <- survey_orgs %>%
  select(home = Q2.2_iso3, target = Q2.5_iso3) %>%
  unnest(target) %>% 
  mutate(across(c(home, target), 
                list(continent = ~countrycode(., "iso3c", "continent",
                                              custom_match = c("XKX" = "Europe"))))) %>% 
  mutate(across(ends_with("_continent"),
                ~recode(., Oceania = "Asia &\nOceania", Asia = "Asia &\nOceania"))) %>% 
  group_by(home_continent, target_continent) %>%
  summarize(times_marked = n()) %>% 
  ungroup() %>% 
  mutate(id = 1:n()) %>% 
  pivot_longer(ends_with("continent"), names_to = "x", values_to = "continent")

ggplot(continents_marked,
       aes(x = x, y = times_marked, stratum = continent, alluvium = id, fill = continent)) +
  geom_flow() +
  geom_stratum(color = "grey30", alpha = 0.5, linewidth = 0.2) +
  geom_text(stat = "stratum", aes(label = continent, color = continent), lineheight = 0.9,
            family = "Fira Sans Semibold") +
  scale_x_discrete(labels = c("Home region", "Target region"),
                   expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(values = c(clrs$Prism[6], clrs$Prism[8], clrs$Prism[11], clrs$Prism[2]),
                    guide = "none") +
  scale_color_manual(values = colorspace::darken(c(clrs$Prism[6], clrs$Prism[8], 
                                                   clrs$Prism[11], clrs$Prism[2]), 0.5),
                    guide = "none") +
  labs(x = NULL, y = NULL,
       title = "Flows between INGO home and target regions",
       subtitle = "European NGOs tend to work in Europe; American NGOs go all over the place") +
  theme_ingo() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(family = "Fira Sans Semibold", size = rel(1.3)),
        axis.text.y = element_blank())
```

## Regions worked in

```{r tab-work-regions}
df_target_regions <- survey_orgs %>%
  select(target = Q2.5_iso3) %>%
  unnest(target) %>% 
  mutate(target = countrycode(target, "iso3c", "continent",
                              custom_match = c("XKX" = "Europe"))) %>% 
  mutate(target = recode(target, Oceania = "Asia &\nOceania", Asia = "Asia &\nOceania")) %>% 
  group_by(target) %>% 
  summarize(times_marked = n()) %>% 
  mutate(prop = times_marked / sum(times_marked)) %>% 
  arrange(times_marked) %>% 
  mutate(target = fct_inorder(target))

df_target_regions %>% 
  gt() %>% 
  cols_label(
    target = "Region",
    times_marked = "Times marked",
    prop = "Proportion"
  ) %>% 
  fmt_percent(columns = prop, decimals = 1) %>% 
  fmt_number(columns = times_marked, decimals = 0) %>% 
  cols_align(align = "left", columns = target) %>% 
  opts_theme() %>% 
  opts_int(use_pagination = FALSE, use_pagination_info = FALSE)
```

```{r plot-work-regions, fig.width=6, fig.height=3}
ggplot(df_target_regions, aes(x = times_marked, y = target)) +
  geom_pointrange(aes(xmin = 0, xmax = times_marked)) +
  scale_x_continuous(labels = label_comma(),
                     sec.axis = sec_axis(~ . / sum(df_target_regions$times_marked),
                                         labels = label_percent())) +
  labs(x = "Times country in region selected", y = NULL,
       title = "Region of target countries",
       subtitle = "Q2.5: Where does your organization work? (summarized by region)") +
  theme_ingo() + 
  theme(panel.grid.major.y = element_blank())
```

# What do these NGOs do?