### 2023-05-10 notebook/desc-details.qmd
Split descriptive stuff across multiple qmd files

---
title: "General respondent details"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "ragg_png")

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(scales)
library(gt)
library(here)

tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(survey_orgs, survey_countries))

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))
```

# How many NGOs responded?

```{r summary-n-respondents}
#| code-fold: show
nrow(survey_orgs)
```

# How many respondents answered questions for more than one country?

```{r summary-n-multiple-countries}
#| code-fold: show
survey_countries %>% 
  summarize(more_than_1 = sum(loop.number > 1),
            prop = more_than_1 / nrow(survey_orgs))
```

# Who in the organization responded to the survey?

```{r plot-respondents-who, fig.width=6, fig.height=3}
df_plot_respondents <- survey_orgs %>%
  group_by(Q2.3) %>%
  summarise(num = n()) %>%
  arrange(num) %>% 
  mutate(question = fct_inorder(Q2.3))

df_plot_respondents %>% 
  ggplot(aes(x = num, y = question)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0),
                     sec.axis = sec_axis(~ . / sum(df_plot_respondents$num),
                                         labels = label_percent())) +
  labs(x = "Respondents", y = NULL,
       title = "Who filled out the survey?",
       subtitle = "Q2.3: What is your position in your organization?") +
  theme_ingo()
```

Lots of others. Who are the others?

```{r tab-who-responded-others}
survey_orgs %>%
  filter(!is.na(Q2.3_TEXT)) %>%
  mutate(`Position in organization` = str_to_title(Q2.3_TEXT)) %>%
  group_by(`Position in organization`) %>%
  summarise(Count = n()) %>%
  mutate(Proportion = Count / sum(Count)) %>%
  arrange(desc(Count)) %>% 
  gt() %>% 
  fmt_percent(columns = Proportion, decimals = 1) %>% 
  opts_int() %>% opts_theme()
```