### 2023-05-12 notebook/desc-governments.qmd
Add relationship with government + issue area

---
title: "INGOs and governments"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "ragg_png")

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(scales)
library(glue)
library(gt)
library(ggmosaic)
library(ggtext)
library(patchwork)
library(tidybayes)
library(here)
library(ggpattern)

tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(survey_orgs, survey_countries, survey_all))
tar_load(c(df_govt_positivity_regime, models_govt_positivity_regime,
           df_govt_positivity_issue, models_govt_positivity_issue))

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))
```

# Relationship with host governments

## Regime type

```{r plot-govt-positivity-regime, warning=FALSE}
ggplot(data = df_govt_positivity_regime) +
  geom_mosaic(aes(x = product(target.regime.type, Q4.11_collapsed), 
                  fill = target.regime.type),
              divider = mosaic("v"), 
              offset = 0, color = "white", linewidth = 1) +
  scale_fill_manual(values = c(clrs$Prism[4], clrs$Prism[6]),
                    guide = "none") +
  labs(x = "Target country regime type",
       y = "Relationship with target country",
       title = glue("How would you characterize your organization’s<br>",
                    "relationship with the government of {country}?",
                    country = build_span("target_country",
                                         c("font-family" = "Consolas", 
                                           "color" = clrs$Prism[1])))) +
  theme_ingo() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_markdown(lineheight = 1.2))
```

```{r plot-govt-positivity-regime-bayes, fig.height=4}
p1 <- models_govt_positivity_regime$draws %>% 
  ggplot(aes(x = .epred, y = Q4.11_collapsed, fill = Q4.11_collapsed)) +
  stat_halfeye() +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = c(clrs$Prism[7], clrs$Prism[12], clrs$Prism[2]),
                    guide = "none") +
  labs(x = "Posterior proportion", y = "Relationship with target country") +
  theme_ingo()

plot_data <- models_govt_positivity_regime$diffs %>% 
  mutate(diffs = factor(Q4.11_collapsed, 
                        levels = c("Positive - Neither", "Neither - Negative", "Positive - Negative")))

diffs_summary <- plot_data %>% 
  group_by(diffs) %>% 
  median_qi(.epred, .width = c(0.5, 0.8, 0.95))

p2 <- models_govt_positivity_regime$diffs %>% 
  mutate(diffs = factor(Q4.11_collapsed, 
                        levels = c("Positive - Neither", "Neither - Negative", "Positive - Negative"))) %>% 
  ggplot(aes(x = .epred, fill = diffs, pattern_fill = diffs)) +
  geom_density_pattern(
    pattern = "stripe",
    pattern_density = 0.5,
    pattern_spacing = 0.2,
    pattern_size = 0,
    trim = TRUE,
    linewidth = 0
  ) +
  geom_pointinterval(data = diffs_summary, aes(x = .epred, xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0, linewidth = 0.5, color = clrs$Prism[8]) +
  facet_wrap(vars(diffs), ncol = 1) +
  scale_fill_manual(values = c(clrs$Prism[2], clrs$Prism[12], clrs$Prism[2]),
                    guide = "none") +
  scale_pattern_fill_manual(values = c(clrs$Prism[12], clrs$Prism[7], clrs$Prism[7]),
                            guide = "none") +
  scale_x_continuous(labels = label_number(scale = 100, style_negative = "minus",
                                           suffix = " pp."),
                     breaks = seq(-0.6, 0.2, by = 0.2)) +
  labs(x = "Percentage point differences", y = NULL) +
  theme_ingo() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())

p1 | p2
```

```{r tab-govt-positivity-regime-bayes}
models_govt_positivity_regime$draws %>% 
  median_qi(.epred) %>% 
  mutate(Q4.11_collapsed = as.character(Q4.11_collapsed),
         type = "Median proportions for NGOs working in autocracies") %>%
  rename(y = .epred, ymin = .lower, ymax = .upper) %>% 
  bind_rows(models_govt_positivity_regime$diffs_summary) %>% 
  mutate(across(c(y, ymin, ymax, p_greater_0), 
                list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(nice_value = glue("{y_nice}<br>({ymin_nice} – {ymax_nice})")) %>% 
  mutate(Q4.11_collapsed = str_replace(Q4.11_collapsed, " - ", " − ")) %>% 
  select(type, Q4.11_collapsed, nice_value, p_greater_0_nice) %>% 
  group_by(type) %>% 
  gt() %>% 
  cols_label(
    Q4.11_collapsed = "",
    nice_value = "Posterior median",
    p_greater_0_nice = "p > 0"
  ) %>% 
  fmt_markdown(nice_value) %>% 
  sub_missing(p_greater_0_nice) %>% 
  tab_footnote(footnote = "95% credible intervals shown in parentheses") %>% 
  opts_theme()
```

## Issue contentiousness

```{r plot-govt-positivity-issue, warning=FALSE}
ggplot(data = df_govt_positivity_issue) +
  geom_mosaic(aes(x = product(potential.contentiousness, Q4.11_collapsed), 
                  fill = potential.contentiousness),
              divider = mosaic("v"), 
              offset = 0, color = "white", linewidth = 1) +
  scale_fill_manual(values = c(clrs$Prism[2], clrs$Prism[8]),
                    guide = "none") +
  labs(x = "Contentiousness of INGO work",
       y = "Relationship with target country",
       title = glue("How would you characterize your organization’s<br>",
                    "relationship with the government of {country}?",
                    country = build_span("target_country",
                                         c("font-family" = "Consolas", 
                                           "color" = clrs$Prism[1])))) +
  theme_ingo() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_markdown(lineheight = 1.2))
```

```{r plot-govt-positivity-issue-bayes, fig.height=4}
p1 <- models_govt_positivity_issue$draws %>% 
  ggplot(aes(x = .epred, y = Q4.11_collapsed, fill = Q4.11_collapsed)) +
  stat_halfeye() +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = c(clrs$Prism[7], clrs$Prism[12], clrs$Prism[2]),
                    guide = "none") +
  labs(x = "Posterior proportion", y = "Relationship with target country") +
  theme_ingo()

plot_data <- models_govt_positivity_issue$diffs %>% 
  mutate(diffs = factor(Q4.11_collapsed, 
                        levels = c("Positive - Neither", "Neither - Negative", "Positive - Negative")))

diffs_summary <- plot_data %>% 
  group_by(diffs) %>% 
  median_qi(.epred, .width = c(0.5, 0.8, 0.95))

p2 <- plot_data %>% 
  ggplot(aes(x = .epred, fill = diffs, pattern_fill = diffs)) +
  geom_density_pattern(
    # aes(y = after_stat(density * 1e5)),
    pattern = "stripe",
    pattern_density = 0.5,
    pattern_spacing = 0.2,
    pattern_size = 0,
    trim = TRUE,
    linewidth = 0
  ) +
  geom_pointinterval(data = diffs_summary, aes(x = .epred, xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0, linewidth = 0.5, color = clrs$Prism[8]) +
  facet_wrap(vars(diffs), ncol = 1) +
  scale_fill_manual(values = c(clrs$Prism[2], clrs$Prism[12], clrs$Prism[2]),
                    guide = "none") +
  scale_pattern_fill_manual(values = c(clrs$Prism[12], clrs$Prism[7], clrs$Prism[7]),
                            guide = "none") +
  scale_x_continuous(labels = label_number(scale = 100, style_negative = "minus",
                                           suffix = "\n pp."),
                     breaks = seq(-0.4, 0.1, by = 0.1)) +
  labs(x = "Percentage point differences", y = NULL) +
  theme_ingo() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())

p1 | p2
```

```{r tab-govt-positivity-issue-bayes}
models_govt_positivity_issue$draws %>% 
  median_qi(.epred) %>% 
  mutate(Q4.11_collapsed = as.character(Q4.11_collapsed),
         type = "Median proportions for NGOs working on contentious issues") %>%
  rename(y = .epred, ymin = .lower, ymax = .upper) %>% 
  bind_rows(models_govt_positivity_issue$diffs_summary) %>% 
  mutate(across(c(y, ymin, ymax, p_greater_0), 
                list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(nice_value = glue("{y_nice}<br>({ymin_nice} – {ymax_nice})")) %>% 
  mutate(Q4.11_collapsed = str_replace(Q4.11_collapsed, " - ", " − ")) %>% 
  select(type, Q4.11_collapsed, nice_value, p_greater_0_nice) %>% 
  group_by(type) %>% 
  gt() %>% 
  cols_label(
    Q4.11_collapsed = "",
    nice_value = "Posterior median",
    p_greater_0_nice = "p > 0"
  ) %>% 
  fmt_markdown(nice_value) %>% 
  sub_missing(p_greater_0_nice) %>% 
  tab_footnote(footnote = "95% credible intervals shown in parentheses") %>% 
  opts_theme()
```

# Experiences with regulations

## Familiarity with regulations

## Frequency of changes

## How NGOs find out about changes

## Are NGOs registered?

# Government involvement in INGO activity

## Contact and reporting frequency