### 2023-12-17 notebook/desc-restrictions.qmd
Start updating restrictions notebook

---
title: "INGOs and restrictions"
format:
  html:
    code-fold: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", fig.retina = 3,
                      fig.width = 6, fig.height = (6 * 0.618),
                      out.width = "80%", collapse = TRUE,
                      dev = "ragg_png")

options(digits = 3, width = 120,
        dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(targets)
library(scales)
library(glue)
library(gt)
library(ggmosaic)
library(ggtext)
library(patchwork)
library(tidybayes)
library(here)
library(ggpattern)
library(countrycode)

tar_config_set(store = here('_targets'),
               script = here('_targets.R'))

tar_load(c(survey_orgs, survey_countries, survey_all, dcjw, gwf))
tar_load(c(
  df_restrictions_regime, models_restrictions_regime,
  df_restrictions_issue, models_restrictions_issue,
  df_restrictions_reg_regime, models_restrictions_reg_regime
))

# Plotting functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))
set_annotation_fonts()
```

# Routine regulation vs. burdensome regulation

```{r plot-dcjw-registration, fig.width=6, fig.height=2, warning=FALSE}
dcjw_registration <- dcjw %>%
  mutate(question = recode(question,
    q_2a = "registration",
    q_2b = "burdensome",
    q_2c = "appeal",
    q_2d = "diff.foreign"
  )) %>%
  filter(question %in% c("registration", "burdensome"))

dcjw_registration_panel <- dcjw_registration %>%
  tidyr::expand(Country, question,
    year = min(.$year, na.rm = TRUE):2013
  ) %>%
  left_join(
    dcjw_registration,
    by = c("Country", "question", "year")
  ) %>%
  # Bring most recent legislation forward
  group_by(Country, question) %>%
  mutate(value = zoo::na.locf(value, na.rm = FALSE)) %>%
  ungroup() %>%
  mutate(value = ifelse(is.na(value), 0, value)) %>%
  mutate(cowcode = countrycode(Country, "country.name", "cown",
    custom_match = c(`Serbia` = 340L)
  ))

regime_types <- gwf %>%
  select(cowcode, year, gwf.ever.autocracy) %>%
  group_by(cowcode) %>%
  summarise(ever.autocracy = any(gwf.ever.autocracy, na.rm = TRUE))

dcjw_registration_summary <- dcjw_registration_panel %>%
  left_join(regime_types, by = "cowcode") %>%
  # There are a few countries that didn't have regime type from full.data
  mutate(ever.autocracy = case_when(
    .$Country %in% c(
      "Bahrain", "Equatorial Guinea", "Kuwait", "Oman",
      "Singapore", "United Arab Emirates"
    ) ~ TRUE,
    .$Country == "Serbia" ~ FALSE,
    TRUE ~ .$ever.autocracy
  )) %>%
  group_by(year, question, ever.autocracy) %>%
  summarise(prop.with.reg = sum(value) / n()) %>%
  ungroup() %>%
  mutate(
    question =
      factor(question,
        levels = c("registration", "burdensome"),
        labels = c(
          "Registration required",
          "Registration burdensome"
        )
      )
  ) %>%
  mutate(
    ever.autocracy =
      factor(ever.autocracy,
        levels = c(TRUE, FALSE),
        labels = c("Autocracies", "Democracies")
      )
  ) %>%
  mutate(ever.autocracy = factor(ever.autocracy,
    levels = rev(levels(ever.autocracy)),
    ordered = TRUE
  )) %>%
  filter(year > 1990)

ggplot(
  dcjw_registration_summary,
  aes(x = year, y = prop.with.reg, colour = ever.autocracy)
) +
  geom_line() +
  labs(x = NULL, y = "Proportion of countries\nwith regulations") +
  scale_y_continuous(labels = label_percent()) +
  coord_cartesian(ylim = c(0, 0.65)) +
  scale_colour_manual(values = c(clrs$Prism[4], clrs$Prism[6]), 
    name = NULL, na.translate = FALSE) +
  theme_ingo() +
  facet_wrap(vars(question))
```

# Overall restriction

## Regime type

```{r plot-restrictions-regime, warning=FALSE}
ggplot(data = df_restrictions_regime) +
  geom_mosaic(aes(x = product(target.regime.type, Q4.17_collapsed), 
                  fill = target.regime.type),
              divider = mosaic("v"), 
              offset = 0, color = "white", linewidth = 1) +
  scale_fill_manual(values = c(clrs$Prism[4], clrs$Prism[6]),
                    guide = "none") +
  labs(x = "Target country regime type",
       y = "Level of restriction in target country") +
  theme_ingo() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_markdown(lineheight = 1.2))
```

```{r plot-restrictions-regime-bayes, fig.width=7.5, fig.height=4}
# Template: 60% of INGOs that reported feeling very restricted work in autocracies
p1 <- models_restrictions_regime$draws %>% 
  ggplot(aes(x = .epred, y = Q4.17_collapsed, fill = Q4.17_collapsed)) +
  stat_halfeye() +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = clrs$Peach[c(1, 3, 5, 7)],
                    guide = "none") +
  labs(x = "Posterior proportion", y = "Level of restriction in target country") +
  facet_wrap(vars("Proportion of INGOs that reported level\nof restriction that work in autocracies")) +
  theme_ingo()

plot_data <- models_restrictions_regime$diffs %>%
  mutate(diffs = factor(Q4.17_collapsed,
    levels = c(
      "Very restricted - Moderately restricted",
      "Moderately restricted - Slightly restricted",
      "Slightly restricted - Not restricted"
    )
  ))

diffs_summary <- plot_data %>% 
  group_by(diffs) %>% 
  median_qi(.epred, .width = c(0.5, 0.8, 0.95))

p2 <- models_restrictions_regime$diffs %>% 
  mutate(diffs = factor(Q4.17_collapsed,
    levels = c(
      "Very restricted - Moderately restricted",
      "Moderately restricted - Slightly restricted",
      "Slightly restricted - Not restricted"
    )
  )) %>% 
  ggplot(aes(x = .epred, fill = diffs, pattern_fill = diffs)) +
  geom_density_pattern(
    pattern = "stripe",
    pattern_density = 0.5,
    pattern_spacing = 0.2,
    pattern_size = 0,
    trim = TRUE,
    linewidth = 0
  ) +
  geom_pointinterval(data = diffs_summary, aes(x = .epred, xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0, linewidth = 0.5, color = clrs$Prism[8]) +
  facet_wrap(vars(diffs), ncol = 1) +
  scale_fill_manual(values = c(clrs$Peach[7], clrs$Peach[5], clrs$Peach[3]),
                    guide = "none") +
  scale_pattern_fill_manual(values = c(clrs$Peach[5], clrs$Peach[3], clrs$Peach[1]),
                            guide = "none") +
  scale_x_continuous(labels = label_number(scale = 100, style_negative = "minus",
                                           suffix = " pp."),
                     breaks = seq(-0.6, 0.2, by = 0.2)) +
  labs(x = "Percentage point differences", y = NULL) +
  theme_ingo() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())

p1 | p2
```

```{r tab-restrictions-regime-bayes}
models_restrictions_regime$draws %>% 
  median_qi(.epred) %>% 
  mutate(Q4.17_collapsed = as.character(Q4.17_collapsed),
         type = "Median proportions for INGOs working in autocracies") %>%
  rename(y = .epred, ymin = .lower, ymax = .upper) %>% 
  bind_rows(models_restrictions_regime$diffs_summary) %>% 
  mutate(across(c(y, ymin, ymax, p_greater_0), 
                list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(nice_value = glue("{y_nice}<br>({ymin_nice} – {ymax_nice})")) %>% 
  mutate(Q4.17_collapsed = str_replace(Q4.17_collapsed, " - ", " − ")) %>% 
  select(type, Q4.17_collapsed, nice_value, p_greater_0_nice) %>% 
  group_by(type) %>% 
  gt() %>% 
  cols_label(
    Q4.17_collapsed = "",
    nice_value = "Posterior median",
    p_greater_0_nice = "p > 0"
  ) %>% 
  fmt_markdown(nice_value) %>% 
  sub_missing(p_greater_0_nice) %>% 
  tab_footnote(footnote = "95% credible intervals shown in parentheses") %>% 
  opts_theme()
```

## Issue contentiousness

```{r plot-restrictions-issue, warning=FALSE}
ggplot(data = df_restrictions_issue) +
  geom_mosaic(aes(x = product(potential.contentiousness, Q4.17_collapsed), 
                  fill = potential.contentiousness),
              divider = mosaic("v"), 
              offset = 0, color = "white", linewidth = 1) +
  scale_fill_manual(values = c(clrs$Prism[2], clrs$Prism[8]),
                    guide = "none") +
  labs(x = "Contentiousness of INGO work",
       y = "Level of restriction in target country") +
  theme_ingo() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_markdown(lineheight = 1.2))
```

```{r plot-restrictions-issue-bayes, fig.width=7.5, fig.height=4}
p1 <- models_restrictions_issue$draws %>% 
  ggplot(aes(x = .epred, y = Q4.17_collapsed, fill = Q4.17_collapsed)) +
  stat_halfeye() +
  scale_x_continuous(labels = label_percent()) +
  scale_fill_manual(values = clrs$Peach[c(1, 3, 5, 7)],
                    guide = "none") +
  labs(x = "Posterior proportion", y = "Level of restriction in target country") +
  facet_wrap(vars("INGOs with contentious issues")) +
  theme_ingo()

plot_data <- models_restrictions_issue$diffs %>% 
  mutate(diffs = factor(Q4.17_collapsed,
    levels = c(
      "Very restricted - Moderately restricted",
      "Moderately restricted - Slightly restricted",
      "Slightly restricted - Not restricted"
    )
  ))

diffs_summary <- plot_data %>% 
  group_by(diffs) %>% 
  median_qi(.epred, .width = c(0.5, 0.8, 0.95))

p2 <- plot_data %>% 
  ggplot(aes(x = .epred, fill = diffs, pattern_fill = diffs)) +
  geom_density_pattern(
    # aes(y = after_stat(density * 1e5)),
    pattern = "stripe",
    pattern_density = 0.5,
    pattern_spacing = 0.2,
    pattern_size = 0,
    trim = TRUE,
    linewidth = 0
  ) +
  geom_pointinterval(data = diffs_summary, aes(x = .epred, xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0, linewidth = 0.5, color = clrs$Prism[8]) +
  facet_wrap(vars(diffs), ncol = 1) +
  scale_fill_manual(values = c(clrs$Peach[7], clrs$Peach[5], clrs$Peach[3]),
                    guide = "none") +
  scale_pattern_fill_manual(values = c(clrs$Peach[5], clrs$Peach[3], clrs$Peach[1]),
                            guide = "none") +
  scale_x_continuous(labels = label_number(scale = 100, style_negative = "minus",
                                           suffix = "\n pp."),
                     breaks = seq(-0.4, 0.4, by = 0.2)) +
  labs(x = "Percentage point differences", y = NULL) +
  theme_ingo() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.y = element_blank())

p1 | p2
```

```{r tab-restrictions-issue-bayes}
models_restrictions_issue$draws %>% 
  median_qi(.epred) %>% 
  mutate(Q4.17_collapsed = as.character(Q4.17_collapsed),
         type = "Median proportions for NGOs working on contentious issues") %>%
  rename(y = .epred, ymin = .lower, ymax = .upper) %>% 
  bind_rows(models_restrictions_issue$diffs_summary) %>% 
  mutate(across(c(y, ymin, ymax, p_greater_0), 
                list(nice = ~label_number(accuracy = 0.01)(.)))) %>% 
  mutate(nice_value = glue("{y_nice}<br>({ymin_nice} – {ymax_nice})")) %>% 
  mutate(Q4.17_collapsed = str_replace(Q4.17_collapsed, " - ", " − ")) %>% 
  select(type, Q4.17_collapsed, nice_value, p_greater_0_nice) %>% 
  group_by(type) %>% 
  gt() %>% 
  cols_label(
    Q4.17_collapsed = "",
    nice_value = "Posterior median",
    p_greater_0_nice = "p > 0"
  ) %>% 
  fmt_markdown(nice_value) %>% 
  sub_missing(p_greater_0_nice) %>% 
  tab_footnote(footnote = "95% credible intervals shown in parentheses") %>% 
  opts_theme()
```

# Registration and restrictions

```{r plot-registration-restrictions, fig.width=6.5, fig.height=2}
ggplot(data = df_restrictions_reg_regime) +
  geom_mosaic(
    aes(
      x = product(regime_registered, Q4.17_collapsed),
      fill = regime_registered
    ),
    divider = mosaic("v"),
    offset = 0, color = "white", linewidth = 1, na.rm = TRUE
  ) +
  scale_x_productlist(
    expand = c(0, 0.02),
    labels = c("Registered", "", "Not registered", "")) +
  scale_y_productlist(expand = c(0, 0.02)) +
  scale_fill_manual(
    values = c(
      clrs$Prism[1], clrs$Prism[11],
      clrs$Prism[7], clrs$Prism[6]
    ),
    guide = "none"
  ) +
  facet_wrap(vars(target.regime.type), ncol = 2) +
  theme_ingo() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

```{r tab-registration-restrictions-diffs}
#| tbl-cap: Median differences in proportion of registered and non-registered INGOs selecting frequency of type of restriction. Probability that the proportion is greater than 0 given in parentheses.
models_restrictions_reg_regime %>% 
  unnest(diffs_summary) %>% 
  mutate(nice_value = glue::glue(
    "{diff}<br>({p})",
    diff = label_number(style_negative = "minus", accuracy = 0.01)(y),
    p = label_number(accuracy = 0.01)(p_greater_0))
  ) %>% 
  ungroup() %>% 
  select(Q4.17_collapsed, target.regime.type, nice_value) %>% 
  pivot_wider(names_from = "target.regime.type", values_from = "nice_value") %>% 
  gt() %>%
  fmt_markdown(columns = 2:3) %>%
  cols_align(align = "left", columns = Q4.17_collapsed) %>%
  cols_label(Q4.17_collapsed = "") %>% 
  tab_footnote(footnote = "Values show the differences in proportion of registered and non-registered INGOs providing each response (not registered proportion − registered proportion)") %>% 
  opts_theme()
```

# Gatekeeping and program capture restrictions

# Reactions to restrictions

## Size

## Other instrumental concerns