### 2023-12-22 manuscript/manuscript.qmd

---
title: "Amicable Contempt"
subtitle: "A Conceptual Framework for Understanding International NGO Behavior in the Era of Closing Civic Space"
short-title: "Amicable Contempt"
author:
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/truthful-tuna>"
abstract: |
  Stuff here

bibliography: references.json
csl: csl/apa.csl
link-citations: true
word-count: "9,878"
---

```{r setup, include=FALSE}
if (is.null(knitr::pandoc_to())) {
  fmt_out <- "interactive"
} else {
  fmt_out <- knitr::pandoc_to()
}

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE,
  warning = FALSE, message = FALSE,
  fig.align = "center", fig.retina = 3,
  fig.width = 6, fig.height = (6 * 0.618),
  out.width = "100%", collapse = TRUE
)

options(
  digits = 3, width = 120,
  dplyr.summarise.inform = FALSE
)
```

```{r libraries-data}
library(tidyverse)
library(scales)
library(tidybayes)
library(marginaleffects)
library(sf)
library(ggtext)
library(ggalluvial)
library(patchwork)
library(glue)
library(janitor)
library(countrycode)
library(gt)
library(targets)

# Generated via random.org
set.seed(8511)

# Targets stuff
tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)

tar_load(c(
  survey_orgs, survey_countries, survey_all,
  world_map
))

# Plotting and modelsummary functions
invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ingo())
set_annotation_fonts()
```

Lots of work on closing space - DPR2015

the legal environment for civil society organizations is becoming increasingly restricted.

In its 2017 report on global civic space, CIVICUS found that only 3% of the world's population lives in open societies with minimal restrictions on associational activity [@CIVICUS:2017], with the majority living in countries with obstructed or repressed civic space (see @fig:civicus-map).

The space for civil society does indeed appear to be closing, but many INGOs have found creative ways to work around legal restrictions to continue their work. For example, since 1993 the Cairo Institute for Human Rights Studies (CIHRS) worked in Egypt as a think tank, research center, and hub for human rights advocacy for the larger Middle East. Between 2012–2014, however, the legal environment for INGOs working Egypt became increasingly restricted and CIHRS faced significant threats and legal pressure to end its research. In response, in December 2014 CIHRS moved its regional and international operations to Tunisia [@CIHRS:2014]. Civic space had closed in Egypt, but CIHRS found a way around the restrictions and continues to engage in research throughout the region.

Other INGOs have made similar adaptations. In 2014, for instance, Venezuela faced a deep economic recession while simultaneously cracking down on civic space and barring international aid agencies from the country. A group of Venezuelan-Americans in Florida founded El Programa de Ayuda Humanitaria para Venezuela (PAHPV) as a response to the growing humanitarian crisis.[^2] Prohibited from setting up an office in Venezuela itself, PAHPV collected medical equipment, supplies, and other donations throughout the southeastern United States. Lacking official legal status in the country, PAHPV initially partnered with domestic Venezuelan NGOs to obtain proper permits for importing goods. Later, when unable to obtain permission, PAHPV volunteers flew to Caracas carrying supplies and cash in their luggage, distributing the donations to local NGOs upon arrival. Civic space in Venezuela was closed to foreigners, but PAHPV and similar organizations still found their way in.

Hundreds of INGOs face similar regulatory dilemmas and must decide how to respond to increasing government oversight of their programs. Civic space is closing around the world, and INGOs are adapting, but not all adapt equally or with the same success.

What determines whether an INGO can avoid becoming captured by the regime and when it can stay actively engaged in its host country? How do

In this article, I provide a theoretical framework for explaining the dynamic relationship between international NGOs and their host country governments.

I theorize that each party is aware that the other both threatens and supports their existence, thus creating a state of *amicable contempt*. Governments see global civil society as a strategic player in their domestic political landscape. International NGOs can help governments maintain by providing services and expertise, and by bestowing international legitimacy and reputation. At the same time, though, INGOs pose a risk to governments. Their missions and goals might conflict with the ruling regime's—for instance, a human rights INGO might condemn a leader who relies on torture to control dissidents. To address this dilemma, governments enact regulations and restrictions that offset the costs and risks of letting INGOs operate in their country while simultaneously capturing the political, social, and other practical benefits of these organizations.

International NGOs understand this contentious relationship with their host countries and work to adapt to the political environment. Parallel to the cost-benefit calculus of their host governments, INGOs seek to balance their need to pursue their work (often in opposition to the regime) against the risk of acting as a political pawn of the regime and propping it up. Rather than passively accepting the regulatory environment imposed on them, INGOs work to either shape that environment or adjust their programs and strategies to maintain access to the country.

Because of the costs involved in confronting and adapting to their regulatory environment, not all INGOs are able to make these adjustments, and organizations often adjust to fit regime preferences or exit the country. I theorize that an INGO's ability to respond to government regulations is determined by two organizational factors: (1) the degree of an INGO's programmatic and instrumental flexibility influences how it responds to gatekeeping regulations, and (2) an INGO's commitment to its core mission, vision and values influences how it responds to program capture regulations.

I explore the testable implications of this theory with data from a global survey of INGOs. This article provides rich descriptions of INGO responses to regulations. While much work has been done on closing civic space, general INGO responses to regulation, and detailed country-specific work, this article provides rich descriptions

In this article, I use rich descriptive data from a global survey of international NGOs to operationalize and define each element of the theory of amicable contempt. I find that INGOs are primarily motivated by their core vision and values, but that they have to balance the pursuit of their missions with instrumental concerns such as fundraising, time, staffing, and collaboration. These concerns both limit and enable INGO activities—without substantial instrumental resources and programmatic flexibility, organizations are unable to carry out their mission, while too much emphasis on resource concerns distracts organizations from their core programming and reduces their effectiveness.

Similar to DeMattee's typology, I don't test any hypotheses or make any causal claims. Rather, I present a conceptual framework for understanding the dynamic relationship between INGOs and host governments and illustrate its mechanisms using survey data.

# Theory

Brief lit review stuff on nonprofit motivations and relationships to regimes
Aseem Prakash + Bloodgood + Tremblay-Boire + Anthony

An INGO's mission and menu of programs will often conflict with the preferences of the regime. Accordingly, an organization must decide how to adapt its programming to fit the requirements and needs of its host country. This decision can be difficult, pitting conflicting organizational values against each other. If the organization is too antagonistic, contentious, or principled, it runs the risk of expulsion, potentially leaving the INGO unable to continue its work in the country. However if the organization is too cooperative, its selection of programs may not be credible or effective as the INGO bends to the government's demands.

I use an adaptation of negotiation theory [@Raiffa:1982; @Mayer:2010] to conceptualize this dynamic relationship between governments and INGOs. I represent the interaction between regimes and INGOs in any given issue area (human rights, development, advocacy, etc.) in simple two-dimensional policy space (see @fig-theory-full). Each actor has a preference for a certain level of advocacy or service provision, marked as solid circles in the figure. Governments set *de jure* policies to prevent INGO activities from going beyond the regime's ideal point, while INGOs will engage in programming that, when possible, fits within the bounds set by the regime. Each actor has a range of flexibility in the implementation of their regulations or their programming, shown in the horizontal lines flanking the ideal points. The space where the two ranges overlap determines both the regime's *de facto* level of restrictions and the INGO's actual mix of programming in that country and constitutes the practical operating environment for the INGO.

```{r}
#| label: fig-theory-full
#| fig-cap: "Programming preferences in two-dimensional policy space"
#| out-width: "100%"
#| include: true

knitr::include_graphics("graphics-manual/out/theory-full.png", auto_pdf = TRUE)
```

A regime's preferences comprise two components: (1) formal *de jure* legislation, represented by the ideal point, and (2) *de facto* implementation of those policies, represented by the negotiation space around the ideal point. Regimes establish a regulatory environment for international civil society that ranges in permissiveness—some countries allow for a large degree of contentious INGO activity, while others are highly restrictive and provide little room for dissonance. In turn, and analogous to the regime's preferences, an INGO's negotiation preferences comprise two components: (1) the organization's normative principles and ideals, represented by the ideal point, (2) the organization's programmatic flexibility in actually implementing those principles, represented by the negotiation space. INGOs have a desire to engage in a certain level of advocacy, generally defined by their mission, vision, and values. INGOs with more instrumental flexibility—that is, a diverse menu of programs, professionalized staff, consistent revenue streams, and collaborative relationships with governments, foundations, and other NGOs—can fine tune their programs to fit the regulatory environments they face, while those with less flexibility are less able to make such adjustments.

The interaction between the actors' negotiation positions introduces two additional dynamics that are crucial for determining how authoritarian regimes regulate INGOs and how well INGOs can respond correspondingly. First, the distance between the two parties ideal points determines the contentiousness of their relationship. INGOs that engage in programming close to the regime's preferences are naturally more cooperative, either providing services that the government needs or advocating for issues that are unlikely to lead to political instability. Accordingly, INGOs that are more cooperative will be *compliant* with the legal limits imposed by the regime. Organizations that pursue more contentious programming further from the regime's ideal point, on the other hand, will have an *antagonistic* relationship and will be more likely to openly resist the restrictions imposed on them.

Second, the size of the two actors' overlapping negotiation spaces determines the practical operating environment for INGOs. This space falls along a continuum of restrictiveness, ranging from *tolerance*, or very few legal or actual restrictions on INGO programs, to *expulsion*, or a complete ban of all INGO activities. The larger the practical operating environment, the more tolerant the regime will be about international NGOs working in their regimes, and vice versa—when there is no overlap, an INGO will no longer be allowed to operate in the country.

```{r}
#| label: tbl-theory-grid
#| include: true
#| tbl-cap: "Possible combinations of INGO/regime relationships and practical 
#|   operating environments"

tibble(
  col1 = c(
    "**Cooperative**", rep("", 2),
    "**Antagonistic**", rep("", 2)
  ),
  Tolerance = c(
    "The INGO's core mission is close to regime preferences and the INGO/regime relationship is cooperative. De jure laws may be stricter than the INGO's ideal preferences, but because there is overlap between de facto regulations and INGO programs, the regime tolerates the INGO's work and the organization operates comfortably in the country.",
    "cooperation-tolerance",
    "*Example*: The Red Cross or Oxfam in most countries",
    "The INGO's core mission is far from regime preferences and the INGO/regime relationship is antagonistic. Because there are substantial associational rights and freedom for advocacy, and because the INGO has a range of possible programming, the regime tolerates the INGO's work and the organization operates comfortably in the country.",
    "antagonism-tolerance",
    "*Example*: Amnesty International in the United States"
  ),
  Expulsion = c(
    "The INGO's core mission is close to regime preferences and the INGO/regime relationship should theoreticaly be cooperative. However, because the de facto civil society regulatory environment is highly constricted, and because the INGO has limited programmatic flexibility, the regime severely limits or expels the organization.",
    "cooperation-expulsion",
    "*Example*: Ford Foundation or Mercy Corps in India after 2015",
    "The INGO's core mission is far from regime preferences and the INGO/regime relationship is antagonistic. Because the de facto civil society regulatory environment is highly constricted, the regime restricts, bans, or expels the organization even if the INGO has substantial programmatic flexibility.",
    "antagonism-expulsion",
    "*Example*: Amnesty International with North Korea"
  )
) |>
  group_by(group = "Nature of relationship") |> 
  gt(id = "theory-table") |>
  tab_spanner(
    label = "Practical operating environment",
    columns = c(Tolerance, Expulsion)
  ) |> 
  cols_label(col1 = "") |>
  fmt_markdown(columns = everything(), rows = c(1, 3, 4, 6)) |>
  text_transform(
    cells_body(columns = c(Tolerance, Expulsion), rows = c(2, 5)),
    fn = \(x) local_image(filename = glue("graphics-manual/out/{x}.png"))
  ) |>
  tab_style(
    style = cell_borders(sides = "all", color = NULL),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_borders(
      sides = "top",
      color = "#D3D3D3",
      weight = px(2)
    ),
    locations = cells_body(columns = everything(), rows = c(4))
  ) |>
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  ) |> 
  tab_style(
    style = cell_fill("#FCF2E6"),
    locations = cells_row_groups()
  ) |> 
  tab_style(
    style = cell_fill("#EFECF4"),
    locations = cells_column_spanners()
  ) |> 
  tab_options(
    quarto.disable_processing = TRUE,
    column_labels.border.top.color = "#ffffff",
    column_labels.font.weight = "bold",
    stub_row_group.border.color = "#ffffff",
    stub_row_group.font.weight = "bold",
    heading.align = "center",
    row_group.as_column = TRUE
  )
```

The relationship between INGOs and regimes is dynamic, and shifts in INGO flexibility and de facto regime civil society regulations can create different practical operating environments, demonstrated in @tbl-theory-grid. For instance, there are two circumstances where INGOs will feel less restricted. First, as seen in the top left quadrant of @tbl-theory-grid, if the distance between the two parties' ideal points is short and the negotiation spaces overlap—as with INGOs providing non-contentious programming like disaster relief, humanitarian services, or education—INGOs will be cooperative and face a tolerant operating environment. Because of preference alignment, these INGOs will not need to rely on their flexibility to maintain access to their host countries.

More contentious organizations with preferences that diverge from the regime's ideal point, on the other hand, can also face a tolerant environment as long as they have a large amount of programmatic flexibility (see bottom left quadrant of @tbl-theory-grid). For example, @Heiss:2019a explores how human rights INGOs advocate for democratization, political reform, and increased respect for human rights by relying on networks of field offices, funded by dozens of private foundations and government grants—resources that give these organizations the flexibility to remain in their host countries and maintain a fairly tolerant operating environment.

Likewise, there are two circumstances where INGOs will feel restricted, or even expelled. Most often, organizations that face restricted operating environments engage in contentious programming and work in countries where regimes have limited de facto regulatory environments, leaving little room for negotiation or overlap (see bottom right quadrant of @tbl-theory-grid). In contrast, INGOs that are naturally less contentious and more prone to compliance can face expulsion if they lack flexibility or it the civil society regulatory environment is too constrained (see top right quadrant of @tbl-theory-grid). For instance, TODO in 2015, less contentious INGOs like Mercy Corps were restricted after India's de facto environment constricted [@Carothers:2015]. No formal laws were changed, but the amount of leeway granted to INGOs shrunk considerably and Modi's government froze assets and cancelled registrations of hundreds of INGOs.

# Data and methods: a global survey of international NGOs

Growing body of work on the government side of the the model, looking at how and why regimes regulate INGOs and why they differentially enforce regulations. Less work has been done on the INGO side… TODO
 There is extensive cross-country research on the relationship between restrictive governments and *domestic* institutions like civil society [@FrantzKendall-Taylor:2014; @Kendall-TaylorFrantz:2014a; @FrantzEzrow:2011], as well as detailed case studies of specific regimes [@Chen:2012; @Teets:2014; @Hildebrandt:2013], and cross-national quantitative work [@DeMattee; @that-one-group-in-europe], but few studies have examined the motivations and strategies of INGOs themselves.

Studying INGO responses to legal regulations—particularly for organizations working in highly restrictive countries—poses a difficult methodological challenge.

One reason for this is an absence of data. INGOs publish a wealth of unstructured data, such as annual reports, budgets, press releases, or calls to advocacy, but few have systematically compiled these data sources into usable datasets. Existing research tends to rely on event data parsed from large corpora of news reports, or presents highly specific in-depth case studies focused on a handful of prominent organizations, resulting in findings and theories that are likely not very generalizable [@BloodgoodSchmitz:2013; @MitchellSchmitz:2014]. While there is utility in amassing thick descriptions of INGO activities, more must be done. @MitchellSchmitz:2014 argue that INGO case studies need to be combined with large, more representative analyses to better understand the organizations that work on specific issue areas as a whole. To this end, I illustrate this framework of INGO behavior using the results from a global survey of international NGOs that work in countries that have imposed NGO regulations.

Complete technical details of the survey are provided in the appendix. I define an INGO as any organization that is not incorporated as a governmental entity or as a profit-seeking private firm that conducts advocacy, provides services, or otherwise works in at least *one* country other than the one it is headquartered in. Of the 17,590 organizations that potentially received an e-mail invitation to participate in the online survey, 537 completed the survey and 104 provided a partial (but usable) response. This yielded a participation rate of 3.8%, which is admittedly low, but comparable to other internet-based organizational surveys [@BaruchHoltom:2008; @HeissKelley:2016]. Respondents were allowed to answer country-specific questions for multiple countries, thus yielding 641 organization-level responses and 659 country-level responses. The survey included 36 open-ended questions, and most respondents answered between 8–16, providing a wealth of qualitative insights into INGO experiences with their host governments. To give voice to these organizations, I quote extensively from these responses with minimal editing or correction. I guaranteed respondents anonymity to encourage candor in responses and to protect privacy. Accordingly, I assign each respondent a random four-digit ID number, and I strip all potentially identifying information from excerpts and quotes from their responses. I also anonymize any country-specific information, replacing references to country names with general geographic regions.

Those who filled out the survey were quite familiar with their organizations' work and had an intimate perspective of their organizations' experiences and strategies. More than half of the respondents indicated that they were the executive director, president, secretary general, CEO, or other high-level officer for their organization, and roughly a third were staff members or program officers.

Respondents span the globe and work on a wide variety of issues and policies. The sample of INGOs have their headquarters spread across 93 different countries (see @fig-hq-countries), with roughly two thirds based in either Europe or the Americas, primarily in the US, UK, and Switzerland (see left panel of @fig-work-hq-regions).

```{r}
#| label: fig-hq-countries
#| fig-width: 6
#| fig-asp: 0.542
#| fig-cap: "Countries where INGOs are headquartered"
#| include: true
# Through lots of trial and error, I found that the best size for this map that
# eliminates extra whitespace on the top/bottom and left/right is 6x3.25, or an
# aspect ratio of 0.542ish

df_hq_countries <- survey_orgs |>
  group_by(Q2.2_iso3) |>
  summarize(num_ngos = n()) |>
  ungroup() |>
  mutate(
    num_ngos_ceiling = ifelse(num_ngos >= 30, 30, num_ngos),
    ngo_presence = ifelse(num_ngos >= 1, "At least one NGO respondent", NA),
    prop = num_ngos / sum(num_ngos, na.rm = TRUE)
  ) |>
  mutate(country_name = countrycode(Q2.2_iso3, "iso3c", "country.name")) |>
  arrange(desc(num_ngos))

map_hq <- world_map |> 
  left_join(df_hq_countries, by = join_by(ISO_A3 == Q2.2_iso3)) |> 
  # Remove Fiji for the sake of plotting
  filter(ISO_A3 != "FJI")

ggplot() +
  geom_sf(
    data = map_hq, aes(fill = num_ngos_ceiling, text = ISO_A3),
    linewidth = 0.1, color = "grey30"
  ) +
  scale_fill_binned(
    low = clrs$Peach[1], high = clrs$Peach[7], na.value = "grey95",
    labels = c(1, 10, 20, "30+"), breaks = c(1, 10, 20, 30),
    name = "Number of NGOs: ",
    guide = guide_colorsteps(
      barwidth = 7, barheight = 0.4,
      title.vjust = 1
    )
  ) +
  coord_sf(crs = "+proj=robin") +
  theme_ingo_map()
```

```{r}
#| label: fig-work-hq-regions
#| include: true
#| fig-width: 4
#| fig-height: 2
#| out-width: "80%"
#| fig-cap: "Regions where INGOs are headquartered vs. regions about which 
#|   INGOs answered questions"

continents_marked <- survey_orgs |>
  select(home = Q2.2_iso3, target = Q2.5_iso3) |>
  unnest(target) |> 
  mutate(across(c(home, target), 
                list(continent = ~countrycode(., "iso3c", "continent",
                                              custom_match = c("XKX" = "Europe"))))) |> 
  mutate(across(ends_with("_continent"),
                ~recode(., Oceania = "Asia & Oceania", Asia = "Asia & Oceania"))) |> 
  group_by(home_continent, target_continent) |>
  summarize(times_marked = n()) |> 
  ungroup() |> 
  mutate(id = 1:n()) |> 
  pivot_longer(ends_with("continent"), names_to = "x", values_to = "continent") |> 
  mutate(continent_total = sum(times_marked), .by = x) |> 
  group_by(x, continent) |> 
  mutate(prop = sum(times_marked) / continent_total) |> 
  ungroup() |> 
  mutate(nice_label = glue(
    '{continent}<br><span style="font-size: 6.5px">{label_percent(accuracy = 1)(prop)}</span>'
  ))

ggplot(
  continents_marked,
  aes(x = x, y = times_marked, stratum = continent, alluvium = id, fill = continent)
) +
  geom_flow() +
  geom_stratum(color = "grey30", alpha = 0.5, linewidth = 0.1) +
  geom_richtext(
    fill = NA, label.color = NA, label.padding = grid::unit(rep(0, 4), "pt"),
    stat = "stratum", aes(label = nice_label, color = continent), 
    size = 2.2, lineheight = 0.9,
    family = "Fira Sans Semibold"
  ) +
  scale_x_discrete(
    labels = c("Home region", "Target region"),
    expand = c(0, 0)
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(
    values = c(clrs$Prism[6], clrs$Prism[8], clrs$Prism[11], clrs$Prism[2]),
    guide = "none"
  ) +
  scale_color_manual(
    values = colorspace::darken(c(
      clrs$Prism[6], clrs$Prism[8],
      clrs$Prism[11], clrs$Prism[2]
    ), 0.5),
    guide = "none"
  ) +
  labs(x = NULL, y = NULL) +
  theme_ingo() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(family = "Fira Sans Semibold", size = rel(0.7)),
    axis.text.y = element_blank()
  )
```

While comparatively few INGOs are headquartered outside the Global North, these organizations have a worldwide reach. In total, INGOs answered questions about their work in 148 different countries. The regional distribution of INGO target countries reflects their broader scope and is almost a mirror image of these organizations' home regions—nearly two-thirds of respondents answered questions about their work in Africa and Asia (see right panel of @fig-work-hq-regions), the most underrepresented regions for NGO headquarters. This disparity could reflect the fact that INGOs from the Global South without a clear English-language internet presence were excluded from the sample, but the pattern also corresponds to similar findings in other surveys of global NGOs [@Limoncelli:2016; @HeissKelley:2016; @FootToftCesare:2015], reinforcing the idea that Northern-based NGOs tend to focus their work in the South. Organizations also operate in a variety of political contexts. Unsurprisingly, given the concentration of INGOs in the Global North, the vast majority of respondents (90%) are headquartered in democracies. The countries these organizations work in, though, show more diversity in regime type—a third of respondents answered questions about their work in autocratic countries.

# Illustration of framework

To illustrate the mechanics of this framework, I use the results from the global survey of INGOs to provide rich descriptive details of how organizations have navigated the tensions between their core organizational principles and their instrumental constraints and programmatic flexibility. I use the survey results to examine the overlap between organizational programmatic flexibility and a regime's regulatory environment, which constitutes the practical operating environment that INGOs must work within.

## Core mission

TODO: mission, vision, values

An organization's foundational mission, vision, and values are represented by the ideal point in the model. What exactly these values are, though, has been the subject of a decades-long debate over the role of NGOs in global politics. International NGOs were initially seen as a "magic bullet" for global governance: highly principled voluntary organizations that promoted human rights, alleviated poverty, provided needed services following disasters, and generally behaved altruistically under the leadership of well-intentioned mangers [@Reimann:2005; @Skjelsbaek:1971; @NyeKeohane:1971; @Bartelson:2006]. However, later research highlighted INGOs' lack of public accountability, potential for corruption, dependence on the financial support of governments they were designed to criticize, and emphasis on pursuing funding over mission [@HulmeEdwards:1997a; @HulmeEdwards:1997b; @Marzouk:1997; @BanksHulmeEdwards:2015; @Reimann:2005].

Building on these critiques, a new trend emerged that treated transnational actors as simultaneously normative and strategic [@Cooley:2002; @Keck:1998; @Kelley:2009a]. @MitchellSchmitz:2014 summarized this debate with their theory of nonprofit principled instrumentalism, or the notion that INGOs must balance their normative mission-driven principles against the instrumental needs of organizational survival. Unlike governments that work toward producing social value (such as maximizing national security or promoting general welfare) and unlike firms that focus on financial performance (such as maximizing profits for shareholders), nonprofit organizations face a tradeoff between their social mission and their need for resources and must pursue both concurrently—they must "instrumentally pursue their principled objectives within the economic constraints and political opportunity structures imposed by their external environments" [@MitchellSchmitz:2014, 489].

INGOs are motivated primarily by their foundational mission and values, and they structure their strategies around this vision to ensure that their programming furthers their mission [@Moore:2000]. Survey respondents confirm this motivation, describing their core mission both in aspirational terms, including "fighting poverty and holding leaders accountable,"[^1193] encouraging "awareness and pro-action towards sustainable development,"[^1267] and promoting "democracy and open society" and "governments' respect for the human rights of children";[^1328_1252] and through more specific statements, such as "improv[ing] healthcare, specifically neurological services, in the least developed nations,"[^1257] or ending "the use of immigration detention around the world."[^1376]

Most respondents report a tight link between programs and missions, stating that "[o]ur mission, values and vision are reflected in everything that we do," that "all our work is geared towards the accomplishment of this objective," and that "our approach matches our mission precisely."[^1245_1267_1272] For many organizations, the "mission and values are the reason we work on these objectives"[^1643]—the normative goals of the organization provide strong motivation to engage in pro-social advocacy and services. Specific programming strategies are also closely linked to INGO missions. For instance, an organization working in South Asia has ambitions to "[b]uild the capacity of local organisations to equip adolescent girls to access their rights and address gender inequality through sport." As expected, much of their programming is centered on sports education and leadership training, but importantly, the organization makes a point that nothing they do is implemented directly: because their core mission is to build local capacity, they ensure that "all the work we do is through local organisations."[^1799] Organizations thus work to ensure that their strategies align with and reflect their core missions, visions, and values.

## Ideal point distance and contentiousness

The distance between the government's preferences for a given issue and an INGO's core mission and vision for that issue determine how contentious their relationship will be, ranging from cooperative to antagonistic. This distance, in turn, influences how governments regulate INGO work. The alignment of organizational missions and strategies with government preferences, therefore, is crucial for understanding government regulation of INGOs. Organizations working on issues that align with regime preferences will be treated more favorably and with less regulatory pressure than INGOs working on more threatening, less aligned issues.

To analyze the favorability of INGO programming to government preferences, I divide the self-reported respondent issues into two levels of potential high or low regime contentiousness based on the potential compatibility of the issue with domestic politics in the host regime [@see Bush:2015 for a similar categorization]. I code programming as highly contentious if it potentially targets government activities that are core to regime survival, thus posing a direct threat to domestic policies and stability. As seen in @fig:issues-worked-on, contentious issues include human rights, advocacy, human trafficking, freedom of expression, conflict, and democracy assistance.

![Issues INGOs work on the most, divided by potential contentiousness](../../Output/figures/3-issues-worked-on){#fig:issues-worked-on}

These highly contentious issues contrast with less contentious topics like education, development, health, and the environment. Governments are more willing to accept intervention and assistance in these areas because doing so poses less of a direct threat to regime stability. Providing a public platform for opposition leaders can empower rival institutions; digging wells typically does not. It is easier for a repressive government to improve its domestic reputation and offset challenges to its political stability by embracing relief organizations, permitting community sport organizations, or partnering with health and environment INGOs.

![Frequency of INGO engagement in different types of activities; organizations answered about one or more activities](../../Output/figures/3-main-activities){#fig:main-activities}

This division of INGOs into contentious and non-contentious programming aligns well with the activities INGOs report engaging in. @fig:main-activities shows how frequently responding organizations work on various types of programming.[^12] If the division of contentiousness is correct, low contention INGOs should provide direct aid and services and engage in research and public education most frequently, since these types of programming pose little threat to the target country regime. High contention INGOs, on the other hand, should deal most often with more threatening activities like advocacy, political mobilization, and government monitoring. There are substantial differences in how frequently high and low contention organizations engage in these types of activities. High contention INGOs are more likely to report always engaging in advocacy, monitoring governments, and mobilizing people than their low contention counterparts, which tend to always provide direct aid.[^13] As seen in the figure, 60% of low contention organizations indicated they almost always provide aid, half almost always engage in research, and less than 40% monitor the government or mobilize people. Interestingly, about half of low contention organizations reported regularly engaging in advocacy. However, this form of advocacy appears to still be non-contentious: respondents describe advocating for "expanded public health programs,"[^14] "agr[i]culture,"[^15] "volunteerism,"[^16] "girls completing secondary education,"[^17] or expanding "access [to] health and justice."[^18]

As expected, high contention organizations are most likely to report always engaging in advocacy (63%) and monitoring government policies (53%), including advocating for the recognition and safety of human rights defenders,[^19]assisting with the "rehabilitation of repentant terrorists,"[^20] promoting religious freedom and human rights, [^21] and "holding … leaders accountable" for their policies and actions.[^22] In aggregate, these more contentious organizations engage in research and public education as the third most common primary activity (53%). The educational activities respondents describe clearly pose more of a threat to the government and are more focused on human rights, such as maintaining a death penalty policy database,[^23] organizing activist training sessions and hackathons,[^24] or developing country-specific media literacy training programs.[^25] These organizations only sometimes provide direct aid and services, in contrast to their low contention counterparts. Finally, both kinds of INGOs are least likely to report always engaging in political mobilization, which is understandable for low contention organizations but surprising for those working on more contentious issues. INGOs may have conflated this kind of work with engaging in research—organizations mentioned political mobilization and campaigns in many of their explanations of their research and public education work.

Though a substantial number of organizations engage in contentious programming, in general, INGOs have a fairly positive relationship with their host governments (@fig:govt-positivity-both), since maintaining some level of government tolerance is essential to staying in the country. Only 10% of respondents report having a somewhat or extremely negative relationship with their host governments, while the remainder either have a positive relationship or are ambivalent.[^26] Opinions of the positivity of this relationship depend on the type of government and the contentiousness of the INGO's issue. 60% of INGOs that reported a negative relationship work in autocracies, while only 20–30% with a positive or ambivalent relationship work in similar regimes.[^27] Similarly, 50% of organizations with a negative relationship work on contentious issues. In fact, 17% of high contention INGOs have a negative relationship with their hosts, in contrast to only 8% of low contention organizations.[^28]

![Opinions of relationship with host government, separated by regime type and issue contentiousness (cell heights represent proportion of responses; cell widths represent proportion of organizations working in regime type or on issue)](../../Output/figures/3-govt-positivity-both){#fig:govt-positivity-both}

Organizations working on low contention issues that align with government needs report having a close and cordial working relationship with their hosts. A health organization working in West Africa explains that "[w]e work very closely with government hospitals and universities, [and we meet with] various health ministers [regularly],"[^29]and a development organization working in Central America reports that "[w]e work very closely with the trade unions, many of whom have representatives in the National Assembly," and that they have a good relationship with the host government's foreign ministry and with their ambassador in the organization's home country.[^30] Government officials are often enthusiastic about NGO work and sometimes "verbally express their gratitude to work with" foreign NGOs.[^31] One education INGO working in a South Asian dictatorship summarizes the experiences of many low-contention organizations optimistically: "We really do not have any problem with them, mater [sic] of fact they work with us so well that when we need any kind of help they are there for us to assist. We have a great relationship."[^32]

Not all organizations enjoy such a warm relationship with the government. Some low contention organizations have poor relationships with their host governments because of poor governance, with one disaster relief INGO working in Southeast Asia reporting that "[t]here is a lot of bureaucracy and corruption. We spend so much time to get approval in the projects. We don't agree to pay bribe[s], which makes or [sic] work even more difficult."[^33] More commonly, though, highly contentious INGOs working in authoritarian countries having a more negative relationship with the regime government activities are often the raison d'être for contentious programming. A human rights INGO in Latin America notes dryly, "most typically, [governments] are the driving force of the human rights abuse we document."[^34] An organization working in Central Africa also identifies their programming as the source of this poor relationship:

> we document and report human rights violation, we organise presse conference to denounce and alert the peoples on pertinent cases, the government is against this and task us to be instrumantalised by europeans and americans, we undergo threats of all forms and some times watch our meetings monitored. [sic][^35]

This organization's mission to report on human rights violations makes its relationship with the government naturally combative, and the government responds by conflating INGO programming with former imperial influence and by closely following the organization's activities. The link between contentious programming and poor relationships is reiterated by other human rights INGOs working in Oceania and the Middle East:

> We are highly critical of them and their policies and laws. We are human rights defenders at risk of prosecution for our work in sharing information with the UN human rights mechanisms. We are monitored by the security agency.[^36]

> The government does not want us to document violations of human rights and fundamental freedoms, including systematic torture and abuse of civilians, civil society activists and women rights defenders.[^37]

The distance between the government and INGO preferences explains the potential contentiousness of the relationship between the two parties—the wider the distance, the more antagonistic the relationship. Results from the survey show that contentious INGOs are more likely to face stricter regulations, increased scrutiny, and worse relationships with their host governments. Interestingly, however, half of contentious INGOs report strong and positive relationships with the government despite their ideological distance. Differences in preferences thus do not fully account for the de facto relationship. For a more complete explanation, we must look at interplay between two other elements of the model: an organization's programmatic flexibility and its overlap with government regulations.

## Programmatic flexibility

According to @MitchellSchmitz:2014, nonprofit organizations are primarily *driven* by their missions and *constrained* by instrumental concerns: worries about money and other resources get in the way of pursuing deeper normative principles. I argue, however, that this view of instrumentality is insufficient. Instrumental concerns, including time, money, or political connections, are not simply constraints. These resources are what allow INGOs to pursue their principled ideals in the first place—they provide organizations with the flexibility to engage in programming and adapt to the political and regulatory environments they face. The region surrounding the organization's ideal point represents its reserve of programmatic flexibility, or the organization's ability and willingness to make changes to its programs in response to regulations.

Flexibility has a dual function. First, it is what allows an INGO to carry out its core mission. Programmatic flexibility is influenced by instrumental concerns, including an organization's size, diversity of funding sources, professionalization of staff, knowledge of the host country—essentially any resource-oriented strategy that permits an organization to implement its mission effectively in dynamic environments. This flexibility is also a function of organization norms—INGOs must be willing to use their instrumental resources to make appropriate adjustments to stay in the country. In some cases, an organization might have substantial income and highly professional staff, but will purposely choose to not engage in advocacy in a restrictive country because of its organizational vision or norms (i.e. its ideal point).

The link between an organization's ideals and its flexibility represents a tension unique to third sector organizations. INGOs must balance their normative principles against the instrumental need of organizational survival. Nonprofit organizations must concern themselves with resources: ignoring necessary overhead concerns in favor of pursuing its vision will limit the organization's ability to effectively fund or carry out that mission (see @Moore:2000, 195). At the same time, the pursuit of flexibility is constrained by nonprofit norms. Dedicating too much time to expanding flexibility through overhead activities like fundraising, professionalization, and marketing can erode public trust in the organization.

Second, programmatic flexibility provides INGOs with a protective margin around their core missions, and their reserve of resources determines how effectively they can respond to different forms of government restriction and regulation. When host regimes constrict the allowable space for INGO activity and put pressure on their ideal points, organizations can respond by relying on their programmatic flexibility to make marginal adjustments to their non-core programming—organizations can marshal new resources, collaborate with more partner organizations, request the support of more powerful external actors, relocate their programs to other countries, or pursue other strategies to circumvent, mitigate, or otherwise reduce the impact of those constraints. The more flexibility an INGO has, the better they can adapt and the more creative and antagonistic they can potentially be; the less flexibility they have, the less so.

The majority of INGOs in the survey maintain a positive relationship with their host countries. This is the "amicable" side of the theory of amicable contempt—organizations can help guarantee their access to the country by acting friendly towards the regime, even if their programming might challenge the regime. But the need for amicability does not mean that organizations must give up on their core vision and mission, taming their programs to completely match their host governments' preferences, nor does it entail that governments will always be friendly to the INGOs they allow in: half of organizations with negative host-country relationships work on contentious issues, but by extension, half of the INGOs at odds with the government engage in less contentious programming. The ability of INGOs to remain in their host countries, avoid restrictions, and maintain positive relationships with the government does not depend entirely on ideal preferences. Both actors can emphasize the "contempt" of amicable contempt by making changes to the space around their ideal points.

The survey reveals multiple sources of programmatic flexibility. Most of the organizations that responded to the survey were small, with half employing eight or fewer staff members. This does not mean, however, that the INGO sector as a whole employs few people—because of dozens of large organizations with hundreds (and even thousands) of employees, the average staff size is 50. These larger, better-staffed organizations inherently need to have access to more financial resources to pay their employees, which entails a greater need for consistent revenue flows. At the same time, having more employees provides these larger organizations with more flexibility in the face of restrictions—if an organization is limited in one country, it can more easily shift programs and staff around to work in other countries or on other topics.

<div id="fig:funding-size">
![Source of funding](../../Output/figures/3-funding-size-totals){#fig:funding-size-totals}

![Reliance on funding sources](../../Output/figures/3-funding-size-pct){#fig:funding-size-pct}

Level of reliance on sources of funding, divided by organization size
</div>

Variation in funding sources also helps determine an organization's flexibility. Government grants—from host countries, target countries, and other donor countries—are the most common stream of income reported by respondents, followed by individual donations, grants from NGOs and foundations, and corporate giving (see @fig:funding-size-totals). Additionally, dozens of respondents report generating their own income through fundraising efforts, member fees, consulting services, and unrelated business income. The size of the organization is related to the funding sources it relies on. Organizations that employ more than 50 people selected an average of 4 streams of funding, while smaller organizations reported an average of 3. The mix and selection of funding sources can also depend on organization size (see @fig:funding-size-pct). Smaller organizations are most likely to rely primarily on fundraising, member fees, and individual donations, and less heavily on grants from NGOs, foundations, and corporations. Larger organizations are far more likely than their smaller counterparts to draw money from governments, IGOs, and corporations. Diversification in income sources provides flexibility—if an organization fails to obtain a government grant, it can turn to foundations, fundraising, or other sources to compensate. Organizations with more sources of funding are thus likely more flexible when facing restrictions on their programs abroad.

![Main obstacles to INGO mission; multiple obstacles allowed](../../Output/figures/3-obstacles){#fig:obstacles}

Flexibility is self-reinforcing. The presence of some forms of flexibility, such as a professionalized staff with employees dedicated to fundraising, make it easier to obtain additional resources [@Fogal:2010], and vice versa—the absence of flexibility imposes a host of challenges to INGO missions. Nearly 70% of respondents identified funding as the primary obstacle to pursuing their mission (see @fig:obstacles), followed by other instrumental obstacles like needing to professionalize staff, struggling to work well with local employees and staff, and coordinating efficiently with other NGOs. Without consistent funding, organizations struggle to hire staff, train volunteers, build relationships with local bureaucrats and politicians, or engage in long-term programs in their host countries. A health advocacy organization working in East Africa notes the dampening effect a lack of funding has on its flexibility, stating that the biggest obstacle to fulfilling its mission is

> [c]ore funding so we can do our work and stop spending so much time on raising money to do our work. We are activists, researchers, communications experts, we don't want to spend our lives raising money, schmoozing funding partners and writing proposals.[^38]

These obstacles and challenges to flexibility shape and limit how well INGOs can pursue their missions, but INGOs do not generally redefine their missions in response to these challenges. Rather, when organizations make strategic adjustments to enhance access to funds, professionalize staff, or improve programmatic flexibility in general, they attempt do so without modifying their core missions. "The overall goal of the organization should remain the same," explains an anti-human trafficking INGO working in Southeast Asia, while "strategies to achieve that goal should change and grow on a regular basis, in order to adapt to the development of trends and new technologies."[^39] Similarly, a development INGO in the Caribbean identified a need to adjust instrumental strategies while maintaining their core mission: "Our goals are sound, and the way we do our work is effective, but I feel we need to revamp our fundraising strategy."[^40] Protecting the core mission is essential in the face of financial constraints, and organizations will engage in creative adjustments to maintain their values. For instance, a development INGO in Latin America relates that "[f]unding is always an issue, but in the last 5 years we have changed our strategy to work as a consultancy firm (*without losing our values, vision and mission*) so we can achieve our annual budget."[^41] Even though it changed its legal organizational form, the organization claims that the shift enabled it to more effectively "improve the business ecosystem" and "create value to society" throughout Latin America.

A human rights INGO working in Latin America summarizes the importance of flexibility succinctly, explaining that the largest obstacle to carrying out their mission is "the small size of the organization and the lack of economic resources to ensure the maintenance of activities independently of the public authorities."[^42] Without robust programmatic flexibility, INGOs risk losing both access to their host countries and independence from interference by their host countries.

Thus, international NGOs are generally driven by their core missions and principles, but must pursue their programming within instrumental constraints. The results from the survey strengthen previous work by @MitchellSchmitz:2014 and others, demonstrating how INGOs find themselves caught in a dilemma between principles and instrumentality. I add to this work by arguing that instrumental concerns are not always constraints—the pursuit of resources enables INGO missions in the first place and provides organizations with programmatic flexibility. Additionally, flexibility allows organizations to protect their core missions and avoid changing them in the face of obstacles. The difficulty for nonprofit organizations is to strike the proper balance—focusing too much on mission and ignoring the need for resources can lead to untenable and unfunded programming, while focusing too much on obtaining resources and enhancing flexibility can become an inferior substitute for an organization's more principled goals.

## Practical operating environment

Even under favorable political circumstances, nonprofit organizations walk a difficult tightrope between the dual pressures of mission and money. This balancing act becomes even more difficult when institutional environments work to constrict, target, and limit their programming [@HeissKelley:2017]. Beyond financial and staffing constraints, a third of respondents indicated that domestic politics, laws, bureaucracies, corruption, and conflict also pose substantial burdens to their programming (see @fig:obstacles). Respondents note that they are limited by "shrinking civil space"[^43] and "increasingly restrictive laws in countries where our members operate"[^44] in the countries they target, by "[c]onstant attempts by public and private actors to restrict freedom of expression and media freedom"[^45] and by direct government interference in their activities.[^46]

Host governments expand and contract the de facto allowable space for civil society advocacy based on concerns for regime stability and other shifts in internal politics, external threats, and reputational pressures. For their part, INGOs must ensure that their activities fall within the bounds set by the regime. The overlap of an organization's programmatic flexibility and a government's de facto regulatory environment determines the practical operating environment for the INGO. NGO regulations are not inherently limiting—many regulations exist to help regularize INGO-state relations and represent a large overlapping space [@BloodgoodTremblay-Boire:2016]. As the size of the overlap decreases, however, the practical operating environment for INGOs becomes constrained and organizations risk expulsion.

Because governments set the initial conditions for operating, keeping track of the regulatory environment is crucial for INGOs to make appropriate adjustments and maintain legal status their host countries. INGOs tend to be quite aware of the regulatory environment of their host countries, with more than 80% of respondents somewhat or very familiar with the laws regulating their activities (see @fig:familiarity-both).[^47] Familiarity is affected by the type of regime organizations work in. More than 75% of respondents who answered that they do not know anything about regulations work in democracies, while those most familiar with the regulatory environment are split roughly equally between democracies and autocracies.[^48] INGOs working in dictatorships are the most likely to be the most familiar with regulations, likely because deep knowledge of the legal environment is more essential to operate safely and effectively. The contentiousness of the issues INGOs work on does not help determine their familiarity—high contention INGOs are not more or less likely to be more aware of the legal environment.[^49]

![Knowledge of regulatory environment in host government, separated by regime type and issue contentiousness (cell heights represent proportion of responses; cell widths represent proportion of organizations working in regime type or on issue)](../../Output/figures/3-familiarity-both){#fig:familiarity-both}

INGOs report that the regulatory environment can change frequently—half of respondents report that regulations and laws change every few years, and nearly 20% say regulations change at least once a year (see @fig:frequency).[^50] The regime type of the host country is closely related to how frequently INGOs see changes in regulations. Nearly two thirds of the INGOs that report changes at least once a year work in autocracies, which contrasts starkly with the two thirds of INGOs that experience a static legal environment in democratic host countries.[^51]

![Frequency of changes in regulatory environment for NGOs in host country, separated by regime type (cell heights represent proportion of responses; cell widths represent proportion of organizations working in regime type)](../../Output/figures/3-frequency){#fig:frequency}

Given their frequency, the regulatory changes INGOs observe do not necessarily correlate with formal statutory developments. These laws give bureaucrats and politicians significant interpretive leeway, however, and allow them to reshape the regulatory environment for international NGOs [@ChaudhryHeiss:2022]. Because of this dynamic discretionary de facto environment, INGOs actively seek out information about changes in laws, rules, and regulations. International NGOs are most likely to hear about new legal changes from other NGOs in the country, both from their local partners and from other foreign organizations. They also receive news about changes directly from government officials, sometimes at inopportune moments. One INGO working in East Africa, unaware of any changes in regulations, discovered new funding requirements for foreign NGOs when requesting permits to expand their operations in the country—the organization ultimately decided to not expand as a result.[^52] INGOs also find out about changes in laws through the media and through internet research, including websites that aggregate information about civil society regulations like Civicus and the International Center for Not-for-Profit Law (ICNL).

Due to increased concerns of regime stability in authoritarian countries [@competitve-authnism], regime type influences the size of the de facto regulatory environment. Following the framework, autocracies should have a narrow allowable space, leading to a smaller overlap or practical operating environment. Nearly two-thirds of the survey respondents that feel the most constrained by government regulations work in autocracies, compared to 25% of organizations that report facing no restrictions in similar regimes (see @fig:restriction-both).[^53] The relationship between host country regime type and feelings of restriction is quite marked, with the proportion of respondents working in autocracies decreasing steadily as the feeling of restriction decreases.

![Feeling of restriction in host country across regime type separated by regime type (cell heights represent proportion of responses; cell widths represent proportion of organizations working in regime type)](../../Output/figures/3-restriction-both){#fig:restriction-both}

The overlap of INGO programmatic flexibility and a government's de facto regulatory environment thus determines the practical operating environment for INGO activities, which in turn imposes additional constraints on these organizations. All nonprofit organizations face an internal conflict between mission and money, but nonprofits working as guests in repressive host countries are subject to regulatory vagaries linked to domestic political struggles. INGOs spend resources to remain aware of changes in the regulatory environment and generally work to ensure that their programs do not lead to expulsion. For most INGOs, maintaining access to the country is critical for successfully undertaking their missions, and as a result, most organizations—including half of the more contentious respondents—try to uphold somewhat of a good relationship with their host governments.

# Discussion and conclusion

International NGOs are motivated by different concerns when contemplating how to react to government restrictions, and institutional and organizational characteristics—namely the contentiousness of the INGO–government relationship and the amount of programmatic flexibility enjoyed by the INGO—help determine these organizations' reactions. Authoritarian governments design their civil society regulatory environment to target different elements of an INGO's organizational preferences—gatekeeping regulations keep potentially harmful organizations out of the country and seek to limit the practical operating environment by reducing organizational flexibility, while program capture regulations allow regimes to take advantage of their guest organizations and seek to distort organizational values. @tbl:ingo-reactions-summary provides a helpful overview of these different pressures, concerns, and reactions.

\stgroup
\footnotesize
\renewcommand\arraystretch{1.6}
!INCLUDE "../../Output/tables/3-ingo-reactions-summary.md"
\fingroup

When confronting gatekeeping regulations, an INGO's primary concern is to gain or maintain access to the country. An INGO's level of contentiousness, in concert with the government's civil society regulatory environment, determines how much regulatory pressure the INGO will face. Highly contentious organizations will face higher scrutiny as regimes attempt to limit their access to the country, head off potential political threats, and make the practical operating environment more difficult. In response, organizations that enjoy greater programmatic flexibility can make creative adjustments on the margins of their missions (i.e. without redefining their core) to adapt to these laws. Organizations with less flexibility, on the other hand, are less likely to be able to adjust to the operating environment and will cease their work, exit the country, avoid entering the country in the first place, or maintain a low profile and avoid contact with the government as much as possible. Correspondingly, because they pose less of a political threat to the regime, low contention organizations will face less pressure and scrutiny and should not need to make as many creative adjustments to remain in the country.

In contrast, when facing regulations designed to capture their programs, INGOs are concerned with protecting their core mission and ensuring that complying with government regulations, cooperating with the host government, and allowing for a degree of control or co-optation does not conflict with their values. While gatekeeping regulations are designed to target all potential organizations and act as a filter, program capture regulations target only those organizations allowed in the country. Organizations with high programmatic flexibility will be able to resist or adjust to government attempts at co-optation, while those with fewer resources and less flexibility will be more likely to either concede to government demands or exit the country. High contention organizations face a different dynamic. INGOs that are able to overcome gatekeeping regulations and *enter* the country are more likely to have high flexibility, and are thus more likely to be able to resist co-optation, while INGOs that remain *outside* of the country and work from exile because of strict gatekeeping laws also are more likely to have high flexibility and can more easily resist co-optation.

---

Can apply to other situations, like looking at regime laws or other stuff like that

[^2]: See [http://www.ayudahumanitariavenezuela.org](http://www.ayudahumanitariavenezuela.org), accessed August 2, 2017.

[^1193]: Response 1193.

[^1267]: Response 1267.

[^1328_1252]: Responses 1328 and 1252.

[^1257]: Response 1257.

[^1376]: Response 1376.

[^1245_1267_1272]: Response 1245, 1267, and 1272.

[^1643]: Response 1643.

[^1799]: Response 1799.

[^12]: "Always" and "Most of the time" responses are collapsed to "Almost always"; "About half the time" and "Sometimes" responses are collapsed to "Sometimes."

[^13]: The probability that the difference in proportion of high and low contention INGOs selecting "Almost always" is greater than zero is generally substantial: high contention organizations = advocacy ($P(\Delta > 0) = 0.01$), monitoring ($P(\Delta > 0) = 0.03$), and mobilizing ($P(\Delta > 0) = 0.13$); low contention organizations = aid and services ($P(\Delta > 0) = 0.87$); there is no substantial difference between the responses of high and low contention organizations for engaging in research and public education ($P(\Delta > 0) = 0.47$). All posterior probabilities are available in @tbl:main-activities-p in the appendix.

[^14]: Response 1385.

[^15]: Response 1605.

[^16]: Response 1472.

[^17]: Response 1419.

[^18]: Response 1448.

[^19]: Response 1339.

[^20]: Response 1233.

[^21]: Responses 1304 and 1703.

[^22]: Response 1193.

[^23]: Response 1558.

[^24]: Responses 1487 and 1660.

[^25]: Response 1643.

[^26]: "Extremely negative" and "Somewhat negative" are collapsed to "Negative"; "Somewhat positive" and "Extremely positive" are collapsed to "Positive."

[^27]: This difference is significant: $P(\% \text{ Negative}_\text{Autocracies} - \% \text{ Positive}_\text{Autocracies} > 0) = 0.97$.

[^28]: This difference is also significant: $P(\% \text{ Negative}_\text{Autocracies} - \% \text{ Positive}_\text{Autocracies} > 0) = 0.97$.

[^29]: Response 1189.

[^30]: Response 1192.

[^31]: Responses 1243, 1253, 1284 and 1323.

[^32]: Response 1507.

[^33]: Response 1224.

[^34]: Response 1375.

[^35]: Response 1339.

[^36]: Response 1408.

[^37]: Response 1458.

[^38]: Response 1211.

[^39]: Response 1194.

[^40]: Response 1574.

[^41]: Response 1225, emphasis added.

[^42]: Response 1188, original in Spanish.

[^43]: Response 1296.

[^44]: Response 1428.

[^45]: Response 1317.

[^46]: Response 1193.

[^47]: It is important to remember that this high percentage is likely tied to who responded to the survey—as explained previously, more than 50% of respondents were the organization's CEO, executive director, or other high-level officer.

 "Extremely familiar" and "Very familiar" are collapsed to "Very familiar"; "Moderately familiar" and "Slightly familiar" are collapsed to "Somewhat familiar."

[^48]: $P(\% \text{ Very familiar}_\text{Autocracies} - \% \text{ Not familiar at all}_\text{Autocracies} > 0) = 0.99$.

[^49]: $P(\% \text{ Very familiar}_\text{High contention} - \% \text{ Not familiar at all}_\text{High contention} > 0) = 0.45$.

[^50]: "Once a month" and "Once a year" are collapsed to "At least once a year."

[^51]: $P(\% \text{ At least once a year}_\text{Autocracies} - \% \text{ Rarely or never}_\text{Autocracies} > 0) = 0.97$.

[^52]: Response 1211.

[^53]: "Very restricted" and "Extremely restricted" are collapsed to "Very restricted."

 $P(\% \text{ Very restricted}_\text{Autocracies} - \% \text{ Not restricted}_\text{Autocracies} > 0) = 1$.

\newpage

# References

::: {#refs}
:::

\newpage

:::: {#appendix-count}

# Appendix

::::